### 2012-11-13 thesis/MA-Thesis.tex
Section 6.2 (Automated constraint generation) and section 6.3.1 (Model checker) are now finished.
The .tex file structure and names have been changed. Now they match the numbers and titles in the table of contents.
have written section 2.3
correction, v.1.
2.3.2 finished
correction update, chapter 4 has a new structure
Almost all correction issues of Andreas and Alwin are realized now (except the ones of Alwin I couldn't read). Chapter 4 and 6 have a new structure.
Final version?

@@

\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{amsmath}

@@

@@

\onehalfspacing
\newpage
\thispagestyle{empty}

@@

\newpage

%select english as language!
\selectlanguage{english}
%\pagenumbering{roman}
\pagenumbering{arabic}
\tableofcontents
\newpage
\listoffigures
\newpage
\listoftables
%\pagenumbering{arabic}

@@

\input{tex/1_abstract}
\input{tex/2_introduction}
\input{tex/3_background}
\input{tex/4_goals}
\input{tex/5_design}
\input{tex/6_implementation}
\input{tex/7_prototypeandevaluation}
\input{tex/8_conclusionandfuturework}

\chapter*{Acknowledgment}

>>>>

%\usepackage[backend=biber,style=unsrt]{biblatex}

@@

\usepackage{tikz}
\usetikzlibrary{arrows, shapes}
\usepackage{pgf-umlsd}
\usepackage{amsmath}

@@

%for bookmark bar
\usepackage[colorlinks,pdfpagelabels,pdfstartview=FitH,bookmarksopen=false,bookmarksnumbered=false,linkcolor=black,plainpages=false,hypertexnames=false,citecolor=black]{hyperref}

\usepackage{algorithm}
\usepackage{algorithmic}

@@

\newpage
\thispagestyle{empty}

@@

\newpage
\onehalfspacing
\selectlanguage{ngerman}
\input{tex/0_zusammenfassung}
\selectlanguage{english}
\input{tex/0_abstract}

\chapter*{Acknowledgment}
%\thispagestyle{empty}

For my master thesis it was my wish to do research abroad in a foreign country. Professor Wolfgang Reif supported me in my plan and managed to give me the unique possibility to visit New Zealand for half an year. Therefore I am deeply grateful to him.

My thesis project was hostet at the University of Auckland, New Zealand, by Professor Bruce MacDonald. I want to thank him for the opportunity to stay and to work at his lab, for all the meetings and discussions we had and for all the support he gave to me.

I also want to thank Felix Kaser, a fellow student of mine, who was with me in New Zealand. During several coffee breaks we frequently discussed ideas for this work and he helped me a lot with his good suggestions.

Chandan Datta, PhD student at the University of Auckland and developer of Robostudio, always offered his help regarding the use of Robostudio or general questions about the healthcare robotics project. He furthermore supported me in getting familiar with the lab and its research topics and with Auckland as a city to live in. Especially in the first time after my arrival in New Zealand I was very grateful for this.

Sincere thanks is also given to Alwin Hoffmann and my supervising tutor Andreas Angerer for assisting me by discussions and hints. They did not only inspect and correct the thesis writing regularly, but they also assisted me in finding the thesis topic in the very beginning and contributed to the results of this work decisively by providing ideas and feedback during the research phase.

Last but not least I want to thank my father Rudolf Nägele for all the effort he put in reading and correcting my work. Since this work relates to different mathematical concepts and formalisms such as LTL, model checking, RBDL, etc., he had to spend plenty of hours in front of the computer in order to get familiar with all these concepts and to check this work for plausibility and correctness.

%\newpage
%\thispagestyle{empty}
%\mbox{}
\newpage

%select english as language!
%\selectlanguage{english}
%\pagenumbering{roman}
%\pagenumbering{arabic}

\pdfbookmark[1]{Inhaltsverzeichnis}{toc}
\tableofcontents
\newpage
\listoffigures
\newpage
%\listoftables
\lstlistoflistings
%\pagenumbering{arabic}

@@

\input{tex/1_introduction}
\input{tex/2_background}
\input{tex/3_goals}
\input{tex/4_visualformalismsforsafetyconstraints}
\input{tex/5_automatedgenerationofsafetyconstraints}
\input{tex/6_theltlcreatorprototype}
\input{tex/7_testscenarioandevaluation}
\input{tex/8_conclusionandfuturework}

### 2012-11-13 thesis/tex/0_abstract.tex
Almost all correction issues of Andreas and Alwin are realized now (except the ones of Alwin I couldn't read). Chapter 4 and 6 have a new structure.
Final version?

Nowadays robotic applications are no longer confined to only industrial use but increasingly assigned to jobs within human workspaces. Even the programming of robots is shifting from experienced software developers to the consumers themselves who need to customize the robot's behaviour in order to match individual requirements.
However, both human robot collaboration as well as the robot's behaviour definition by non experts may be a risk to humans within the robots' environment.
Thus, special thinking is needed about how to cope with safety requirements for robot behaviour specification by consumers.

@@

A graphical editor has been developed which demonstrates both visual constraint creation functionality as well as automated constraint generation.

>>>>

Nowadays robotic applications are no longer confined to only industrial use but increasingly assigned to tasks within human workspaces and living environments. Even the programming of robots is shifting from experienced software developers to the consumers themselves who need to customize the robot's behaviour in order to match individual requirements.
However, both human robot collaboration as well as the robot's behaviour definition by non experts may be a risk to humans within the robots' environment.
Thus, special thinking is needed about how to cope with safety requirements for robot behaviour specification by consumers.

@@

A graphical editor has been developed which integrates and demonstrates both visual constraint creation functionality as well as automated constraint generation and makes them usable in an uniform manner.

### 2012-11-13 thesis/tex/1_introduction.tex
Final version?

Safety critical robot applications require extensive testing or formal verification in order to achieve adequate safe and predictable behaviour.
Even if a program does not control actuators itself and thus cannot directly cause any damage, its output might still be a trigger for dangerous actions involving several actors.
Imagine a medication reminder application running on a service robot which guides a person through the process of medication intake, as investigated in the healthcare robotics project at the
\begin{wrapfigure}{r}{0.58\textwidth}
  \centering
  \includegraphics[width=0.56\textwidth]{cafero_being_used}
  \caption{Service robot being used by older person~\cite{robostudio}.}
  \label{fig:cafero_being_used}
\end{wrapfigure}
University of Auckland, New Zealand \cite{jayawardena12:_desig}. All interaction is guided by showing textual instructions on a display, and by speech generation. An error in the program logic could cause an already taken medication to be reminded again and may injure a patient with Alzheimer's disease by causing a medication overdose.
Scientists are working on applications such as the medication reminder~\cite{p11:_feasib} which run on mobile robots (so called healthbots) serving in nursing homes for the elderly. Equipped with a touch display and several external devices such as a blood pressure measurement tool the robots aim to support and entertain the elderly in their daily activities~\cite{jayawardena12:_desig,5649910}.

@@

The applications for the healthcare robots mentioned above are intended to be developed by healthcare professionals using Robostudio~\cite{robostudio}, a visual programming environment for rapid authoring and customization of complex robot services.
%state machine based processes and a graphical tool for the designing of human robot interaction.
Generally these people don't have profound knowledge of the complex mathematical syntax used for formal methods which allow software behaviour verification.
%temporal logic such as computation tree logic (CTL) or linear temporal logic (LTL).
Nevertheless, certain safety guarantees must somehow be delivered in order to gain users' trust and to prevent any harm or injuries by the healthcare robots.
In case of the medication reminder mentioned above it might be important to ensure that notifications are sent to staff members whenever medication is not taken by the patient, or to avoid duplicate reminders for medication which has already been taken.
A mechanism like this is needed for the specification of such safety constraints; one that does not require special or expert knowledge and is thus easy to use for all kind of programmers. A visual language intended to fill this gap is presented in chapter~\ref{chap:visualformalismsforsafetyconstraints}.

Unfortunately, supplying an easy-to-use editor for constraint creation does not guarantee adequate thinking about sufficent constraints by the developer. Unmotivated or just unexperienced developers may miss important constraints needed in order to achieve a certain safety level.
To support the user in finding significant constraints it might be useful to automatically compute constraint suggestions which make true testimonies about the current designed program. First of all the generated constraints provide an opportunity for the developer to identify reasonable constraints as well as contradictions between constraints and specification. The latter would mean there are errors in the program which lead to undesired constraints. Furthermore, once constraints are created and checked for sanity they can be validated after every program change and thus ensure integrity during the development process. Chapter~\ref{chap:automatedgenerationofsafetyconstraints} introduces a heuristics for finding reasonable safety constraints based on state machine definitions that describe robots' behaviours.

>>>>

Safety critical robot applications require extensive testing or formal verification in order to achieve adequate safe and predictable behaviour.
Even if a program does not control actuators itself and thus cannot directly cause any damage, its output might still be a trigger for dangerous actions involving several actors.
Imagine a medication reminder application running on a service robot which guides a person through the process of medication intake, as investigated in the healthcare robotics project at the
\begin{wrapfigure}{r}{0.58\textwidth}
  \centering
  \includegraphics[width=0.56\textwidth]{cafero_being_used}
  \caption{Service robot being used by older person~\cite{robostudio}.}
  \label{fig:cafero_being_used}
\end{wrapfigure}
Uni\-ver\-si\-ty of Auckland, New Zealand \cite{jayawardena12:_desig}. All interaction is gui\-ded by showing textual instructions on a display, and by speech generation. An error in the program logic could cause an already taken medication to be reminded again and may injure a patient with Alzheimer's disease by causing a medication overdose.
Scientists are working on applications such as the medication reminder~\cite{p11:_feasib} which run on mobile robots (so called healthbots) serving in nursing homes for the elderly. Equipped with a touch display and several external devices such as a blood pressure measurement tool the robots aim to support and entertain the elderly in their daily activities~\cite{jayawardena12:_desig,5649910}.

@@

The applications for the healthcare robots mentioned above are intended to be developed by healthcare professionals using Robostudio~\cite{robostudio}, a visual programming environment for rapid authoring and customization of complex robot services.
%state machine based processes and a graphical tool for the designing of human robot interaction.
Generally these people do not have profound knowledge of the complex mathematical syntax used for formal methods which allow software behaviour verification.
%temporal logic such as computation tree logic (CTL) or linear temporal logic (LTL).
Nevertheless, certain safety guarantees must somehow be delivered in order to gain users' trust and to prevent any harm or injuries by the healthcare robots.
In case of the medication reminder mentioned above it might be important to ensure that notifications are sent to staff members whenever medication is not taken by the patient, or to avoid duplicate reminders for medication which has already been taken.
A mechanism like this is needed for the specification of such safety constraints; one that does not require special or expert knowledge and is thus easy to use for all kind of programmers. Within the scope of this work a visual language has been developed intending to fill this gap.

Unfortunately, supplying an easy-to-use editor for constraint creation does not guarantee adequate thinking about sufficent constraints by the developer. Unmotivated or just unexperienced developers may miss important constraints needed in order to achieve a certain safety level.
To support the user in finding significant constraints it might be useful to automatically compute constraint suggestions which make true testimonies about the current designed program. First of all the generated constraints provide an opportunity for the developer to identify reasonable constraints as well as contradictions between constraints and specification. The latter would mean there are errors in the program which lead to undesired constraints. Furthermore, once constraints are created and checked for sanity they can be validated after every program change and thus ensure integrity during the entire development process. This work introduces a new heuristics for finding such reasonable safety constraints based on state machine definitions that describe robots' behaviours.
